#5  Dimensionality Reduction Techniques

##5.1. Human Data from United Nations 


```{r}
#load libraries
library(GGally)
library(corrplot)
library(tidyr)
library(dplyr)
```


```{r}
```

```{r}
human <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", header = TRUE)
dim(human)
str(human)
```
*Data Frame with 155 Rows and 8 Columns
*The human dataset contains various indicators of the well-being of various countries. The summary shows, there are altogether 155 observations (i.e. countries) and these are the variables:
*Edu2.FM: the ratio of females to males in secondary education
*Labo.FM: the ratio of females to males in the labour force
*Edu.Exp: expected number of years spent in schooling
*Life.Exp: life expectancy in years
*GNI: gross national income
*Mat.Mor: the relativised number of mothers who die at child birth
*Ado.Birth: the rate of teenage pregnancies leading to child birth
*Parli.F: the percentage of female parliamentarians



##5.2. Data Exploration 
```{r}
#p <- ggpairs(human, mapping = aes(col="blue", alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
ggpairs(human)
```
# compute the correlation matrix and visualize it with corrplot
#cor(human) %>% round(digits=2) %>% corrplot.mixed(tl.cex = 0.7,number.cex=0.7,order="hclust",addrect=2)
```{r}
cor_matrix<-cor(human)%>% round(digits =2)
corrplot(cor_matrix, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6)
```
* Edu.Exp is normally distributed
* Edu.Exp is negatively correlated with Ado.Birth 
* Life.Exp is highly correlated to GNI and Edu.Exp 
* Life.Exp is negatively correlated to Mat.Mor
* Parli.F doesnt have much impact on the other variables but correlaetd to Female to Male labour ratio

##5.3. Principal Component Analysis on the not standardized human data

```{r}

pca_human <- prcomp(human)
biplot(pca_human, choices = 1:2, cex = c(0.5,0.5), col = c("orange", "purple"), main="Biplot of the first two principal components for the unscaled data")
```
*Only GNI is visible
*Correlation is negligible
*High variance
*Qatar seems to be outlier

##5.4 Principal Component Analysis on the Standardized human data

**Lets scale the data for better interpretation

```{r}
human_scale <- scale(human)
summary(human_scale)
summary(human)
```

**Check PCA after scaling

```{r}
pca_human <- prcomp(human_scale)
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("orange", "purple"), main="Biplot of the first two principal components for the scaled data")
```

Distribution of Countries explains the individual records. Data is now normalized with different parameters. Countries are now separated with respect to correlations with Principal Components

##5.5. PCA Observations

**Observations
* PC1 has high correlation with Mat.Mor
* PC2 has high correlation with Parli.F
* High correlation between Mat.Mor, Ado.Birth, Edu.Exp, Edu2.FM, Life.Exp and GNi, 
* Life.Exp, GNI, Edu.Exp, Edu2.FM is opposite side of Mat.Mor and Ado.Birth which indicates importance of Edu. Exp and Life expectancy
* Female representation in Parliament seems to indicate high GNI, Life Expectancy and greater Female to Male labor ratio.


##5.6. Multiple Correspondence Analysis (MCA)

In the case of categorical variables dimensionality reduction can be provided using correspondence analysis (CA, n=2 variables) or multiple corresponding analysis (MCA, n>2). MCA is a generalization of PCA and an extension of CA. In this analysis cross tabulation matrices between all variables in the data set are used. The information from cross-tabulations has to be presented in clear graphical form. MCA could be used for example as a pre-processing for clustering.


###5.4.1 The data

Here we use tea data from FactoMineR package.
```{r}
library(FactoMineR)
library(tidyr)
library(ggplot2)
library(dplyr)
data("tea")
keep_columns <- c("Tea","price","How", "how", "sugar", "where","lunch")
tea_time <- dplyr::select(tea, one_of(keep_columns))
summary(tea_time)
```

```{r}
str(tea_time)
```
Selected Parameters : Tea, price, How, how, sugar, where and lunch.

Graphically the date are presented as barplots using ggplot() function.
```{r}
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()+theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```
* Very Few peope drink Cheap tea.
* Most people drink tea alone with tea bags and preferred tea variety is Earl Grey at Chain Shops.

Next MCA analysis on the chosen tea data is provided.
```{r}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
```

```{r}
plot(mca, invisible=c("ind"), habillage = "quali")
```

* Tea Shop usually has unpackaged tea with high prices (p_upscale)
* People usually have Earl Gray and more likely to have it with sugar and milk. 
* Green Tea is cut from all..(Explains a lot)

